Provisional Title

### 2021-05-10 draft/draft.tex
more results and figS4 added --am I getting overexcited with the interpretation of the ressults?

\subsubsection*{Top-dow co-selection is weak}

If communities being coalesced were highly cohesive from the top-down,
the dominant species would co-select the rarer members of its community during coalescence
(\figref[e]{fig1}, left panels).
In this scenario, we would expect the outcome of community coalescence to be predicted by
which of the two dominants is most competitive in pairwise competition.
Coalescence outcomes can be quantified by the similarity between the coalesced and the
invasive communities
(\methodsref{metrics}).
To test this hypothesis, we performed all pairwise competitions between dominant species
in glutamine and citrate environments by mixing them 1:1 on their native media and propagating
the cultures for seven serial transfers, roughly 42 generations
(\methodsref{competitions}).
We then performed all possible pairwise community coalescence experiments by mixing
equal volumes of the communities and propagating the resulting cultures for seven extra
transfers
(\figref[f]{fig1}).
The frequencies of all species in both community-community and dominant-dominant competitions
were determined by 16S illumina sequencing
(\methodsref{sequencing}).
We found that the pairwise competitive ability of an invasive dominant
is only weakly predictive of the performance of the invasive community in coalescence,
as quantified by the relative Bray-Curtis similarity between the coalesced and invasive communities
(\figref[a]{fig2}, $R^2=0.15$, $p<$ for glutamine and $R^2=0.15$, $p<$ for citrate).
Alternative quantifications of community distance also yield weak effects,
and even more so when the metric used
accounts only for the presence/absence of specific species and not for their relative abundance in
the communities
(\figref{figS2}).

@@

\subsection*{Stabilization of environmental communities in simple synthetic environments}
\label{methods:community-assembly}

Communities were stabilized \textit{ex situ} as described in \cite{Goldford2018}.
In short, environmental samples (soil, leaves...) within one meter radius in eight different
geographical locations were collected with sterile
tweezers or spatulas into 50mL sterile tubes (Fig. \mr).
One gram of each sample was allowed to
sit at room temperature in 10mL of phosphate buffered saline (1$\times$PBS) containing
200$\upmu$g/mL cycloheximide to suppress eukaryotic growth.
After 48h, samples were mixed 1:1 with 80\% glycerol and kept frozen at $-80^\circ$C.
Starting microbial communities were prepared by scrapping the frozen stocks into
$200\upmu$L of 1$\times$PBS and adding a volume of $4\upmu$L to $500\upmu$L
of synthetic minimal media (1$\times$M9) supplemented with $200\upmu$g/mL cycloheximide
and 0.07 C-mol/L glutamine or sodium citrate as the carbon source in 96 deep-well plates
(1.2mL; VWR).
Cultures were then incubated still at $30^\circ$C to allow for re-growth.
After 48h, samples were fully homogenized and biomass increase was followed by measuring
the optical density (620nm) of $100\upmu$L of the cultures in a Multiskan FC plate reader
(Thermo Scientific).
Communities were stabilized \cite{Goldford2018} by passaging $4\upmu$L of the cultures into
$500\upmu$L of fresh media (1$\times$M9 with the carbon source)  every 48h for a total of
12 transfers at a dilution factor of 1:100,
roughly equivalent to 80 generations per culture (Fig. \mr).
Cycloheximide was not added to the media after the first two transfers.

\subsection*{Isolation of dominant species}\label{methods:dominants}

For each community, the most abundant colony morphotype at the end of the ninth transfer
was selected, resuspended in $100\upmu$L 1$\times$PBS and serially diluted (1:10).
Next, $20\upmu$L of the cells diluted to $10^{-6}$ were plated in the corresponding synthetic
minimal media and allowed to regrow at $30^\circ$C for 48h. Dominants were then inoculated
into $500\upmu$L of fresh media and incubated still at $30^\circ$C for 48h.
After this period, the communities stabilized for eleven transfers and the isolated dominants
were ready for the competition experiments (Fig \mr) at the onset of the twelfth transfer.

\subsection*{Dominant-dominant and community-community competitions}
\label{methods:competitions}

All possible pairwise dominant-dominant and community-community
competition experiments
were performed by mixing equal volumes ($4\upmu$L) of each of the eight
communities or eight dominants at the onset of the twelfth transfer.
Competitions were set up in their native media,
i.e. in $500\upmu$L of 1$\times$M9 supplemented with 0.07 C-mol/L of
either glutamine or citrate in 96 deep-well plates.
Plates were incubated at $30^\circ$C for 48h.
Pairwise competitions were further propagated for seven serial transfers
(roughly 42 generations; Fig. \mr) by transferring $8\upmu$L of
each culture to fresh media ($500\upmu$L).

\subsection*{Determination of community composition by 16S sequencing}
\label{methods:sequencing}

@@

\subsection*{Metrics of community distance}
\label{methods:metrics}

@@

\subsection*{Simulations}\label{methods:sim}

@@

\begin{figure}[!h]
\centering
\internallinenumbers
\includegraphics[width=12.5cm,keepaspectratio]{figs/figS2.pdf}
\caption{\textbf{Alternative metrics of community distance.}
Quantifying coalescence outcomes using different metrics of community similarity
(\methodsref{metrics}) gives similar results to those shown in \figref[a]{fig2}.
Metrics that account for the relative species abundances (\textbf{a.} Jensen-Shannon similarity or
\figref[a]{fig2} Bray-Curtis similarity) yield higher correlations than less quantitative metrics that only
account for species presence/absence (\textbf{b.} Jaccard similarity or \textbf{c.} Fraction of endemic
invasive species persisting in the coalesced community).}
\label{figS2}
\end{figure}

@@

\begin{figure}[!h]
\centering
\internallinenumbers
\includegraphics[width=10cm,keepaspectratio]{figs/figS3.pdf}
\caption{\textbf{Dominant species have limited effects on coalescence outcomes.}
We repeated the analyses shown in \figref[a]{fig2} and \figref{figS2}, but this time we removed the
dominants from the compositional data prior to quantifying community distances.
The trends observed before are maintained.}
\label{figS3}
\end{figure}

>>>>

\subsubsection*{The strength of top-down co-selection is dependent on the primary supplied resource}

If communities being coalesced were highly cohesive from the top-down,
the dominant species would co-select the rarer members of its community during coalescence
(\figref[e]{fig1}, left panels).
In this scenario, we would expect the outcome of community coalescence to be predicted by
which of the two dominants is most competitive in pairwise competition.
Coalescence outcomes can be quantified by the similarity between the coalesced and the
invasive communities
(\methodsref{metrics}).
To test this hypothesis, we performed all pairwise competitions between dominant species
in glutamine and citrate environments by mixing them 1:1 on their native media and propagating
the cultures for seven serial transfers, roughly 42 generations
(\methodsref{competitions}).
We then performed all possible pairwise community coalescence experiments by mixing
equal volumes of the communities and propagating the resulting cultures for seven extra
transfers
(\figref[f]{fig1}).
The frequencies of all species in both community-community and dominant-dominant competitions
were determined by 16S illumina sequencing
(\methodsref{sequencing}).

We found that, for communities assembled in the glutamine environment,
the pairwise competitive ability of an invasive dominant
is only weakly predictive of the performance of the invasive community in coalescence
(as quantified by the relative Bray-Curtis similarity between the coalesced and invasive communities),
but this correlation was stronger for the citrate communities
(\figref[a]{fig2}, $R^2=0.15$, $p<0.05$ for glutamine and $R^2=0.57$, $p<10^{-4}$ for citrate).
Alternative quantifications of community distance yield similar results,
with weaker effects when the metric used
accounts only for the presence/absence of specific species and not for their relative abundance in
the communities
(\figref{figS2}).
All these metrics include the presence of the dominant species themselves. To better disentangle the effect
that these dominants have on the other member of their communities, we repeated the analysis this time
excluding the dominant species from the compositional data, finding that our results still hold
(\figref{figS3}).
Furthermore, we observed that, in the glutamine communities,
the relative frequency of a dominant against another in head-to-head pairwise competition
is barely predictive of its relative frequency against that same other dominant when the
cohorts are present too, i.e. after coalescence
($R^2=0.04$, $p>0.05$, \figref{figS4}).
On the other hand, pairwise competition between dominants of the citrate communities
is highly correlated with the relative abundance of said dominants in community coalescence
 ($R^2=0.83$, $p<10^{-8}$, \figref{figS4}).

Together, these results suggest that the strength of top-down co-selection depends on
the primary resource supplied to the coalesced communities.
Communities assembled with citrate as the primary supplied resource
display a strong degree of top-down co-selection,
with the fates of the sub-dominant species determined to a large extent by dominant-dominant
pairwise competition. This competition is, in turn, only weakly affected by the presence of the cohorts.
For glutamine communities, although some degree of top-down co-selection is consistent with our data,
the cohorts do not appear to be passively responding to their dominants.
Instead, evidence suggests that the cohorts might be playing an active role in community coalescence.
This finding led us to investigate the potential role of bottom-up ecological co-selection
(\figref[e]{fig1}, right panels), i.e. whether in some cases the dominants may be co-selected
for or against by their cohorts.

@@

\subsubsection*{Stabilization of environmental communities in simple synthetic environments}
\label{methods:community-assembly}

Communities were stabilized \textit{ex situ} as described in \cite{Goldford2018}.
In short, environmental samples (soil, leaves...) within one meter radius in eight different
geographical locations were collected with sterile
tweezers or spatulas into 50mL sterile tubes (\figref[a]{fig1}).
One gram of each sample was allowed to
sit at room temperature in 10mL of phosphate buffered saline (1$\times$PBS) containing
200$\upmu$g/mL cycloheximide to suppress eukaryotic growth.
After 48h, samples were mixed 1:1 with 80\% glycerol and kept frozen at $-80^\circ$C.
Starting microbial communities were prepared by scrapping the frozen stocks into
$200\upmu$L of 1$\times$PBS and adding a volume of $4\upmu$L to $500\upmu$L
of synthetic minimal media (1$\times$M9) supplemented with $200\upmu$g/mL cycloheximide
and 0.07 C-mol/L glutamine or sodium citrate as the carbon source in 96 deep-well plates
(1.2mL; VWR).
Cultures were then incubated still at $30^\circ$C to allow for re-growth.
After 48h, samples were fully homogenized and biomass increase was followed by measuring
the optical density (620nm) of $100\upmu$L of the cultures in a Multiskan FC plate reader
(Thermo Scientific).
Communities were stabilized \cite{Goldford2018} by passaging $4\upmu$L of the cultures into
$500\upmu$L of fresh media (1$\times$M9 with the carbon source)  every 48h for a total of
12 transfers at a dilution factor of 1:100,
roughly equivalent to 80 generations per culture (\figref[b]{fig1}).
Cycloheximide was not added to the media after the first two transfers.

\subsubsection*{Isolation of dominant species}\label{methods:dominants}

For each community, the most abundant colony morphotype at the end of the ninth transfer
was selected, resuspended in $100\upmu$L 1$\times$PBS and serially diluted (1:10).
Next, $20\upmu$L of the cells diluted to $10^{-6}$ were plated in the corresponding synthetic
minimal media and allowed to regrow at $30^\circ$C for 48h. Dominants were then identified
(\figref[c]{fig1}),
inoculated into $500\upmu$L of fresh media and incubated still at $30^\circ$C for 48h.
After this period, the communities stabilized for eleven transfers and the isolated dominants
were ready for the competition experiments at the onset of the twelfth transfer.

\subsubsection*{Dominant-dominant and community-community competitions}
\label{methods:competitions}

All possible pairwise dominant-dominant and community-community
competition experiments
were performed by mixing equal volumes ($4\upmu$L) of each of the eight
communities or eight dominants at the onset of the twelfth transfer.
Competitions were set up in their native media,
i.e. in $500\upmu$L of 1$\times$M9 supplemented with 0.07 C-mol/L of
either glutamine or citrate in 96 deep-well plates.
Plates were incubated at $30^\circ$C for 48h.
Pairwise competitions were further propagated for seven serial transfers
(roughly 42 generations, \figref[f]{fig1}) by transferring $8\upmu$L of
each culture to fresh media ($500\upmu$L).

\subsubsection*{Determination of community composition by 16S sequencing}
\label{methods:sequencing}

@@

\subsubsection*{Metrics of community distance}
\label{methods:metrics}

@@

\subsubsection*{Simulations}\label{methods:sim}

@@

\begin{figure}[!h]
\centering
\internallinenumbers
\includegraphics[width=12.5cm,keepaspectratio]{figs/figS2.pdf}
\caption{\textbf{Alternative metrics of community distance.}
Quantifying coalescence outcomes using different metrics of community similarity
(\methodsref{metrics}) gives similar results to those shown in \figref[a]{fig2}.
Metrics that account for the relative species abundances (Bray-Curtis or Jensen-Shannon similarities)
yield higher correlations than less quantitative metrics that only
account for species presence/absence (Jaccard similarity or the fraction of endemic
invasive species persisting in the coalesced community).
\textbf{a.} Relative Jensen-Shannon similarity 
($R^2=0.15$, $p<0.05$ for glutamine and $R^2=0.53$, $p<5\times10^{-4}$ for citrate)
\textbf{b.} Relative Jaccard similarity
($R^2=0.08$, $p>0.05$ for glutamine and $R^2=0.13$, $p>0.05$ for citrate)
\textbf{c.} Relative survival of invasive endemic species after coalescence
($R^2=0.16$, $p<0.05$ for glutamine and $R^2=0.04$, $p>0.05$ for citrate).}
\label{figS2}
\end{figure}

@@

\begin{figure}[!h]
\centering
\internallinenumbers
\includegraphics[width=10cm,keepaspectratio]{figs/figS3.pdf}
\caption{\textbf{Dominant species have limited effects on coalescence outcomes quantification.}
We repeated the analyses shown in \figref[a]{fig2} and \figref{figS2}, but this time we removed the
dominants from the compositional data prior to quantifying community distances.
The trends observed before are maintained.
\textbf{a.} Relative Bray-Curtis similarity
($R^2=0.20$, $p<0.01$ for glutamine and $R^2=0.34$, $p<0.005$ for citrate)
\textbf{b.} Relative Jensen-Shannon similarity 
($R^2=0.24$, $p<0.005$ for glutamine and $R^2=0.36$, $p<0.005$ for citrate)
\textbf{c.} Relative Jaccard similarity
($R^2=0.09$, $p>0.05$ for glutamine and $R^2=0.11$, $p>0.05$ for citrate)
\textbf{d.} Relative survival of invasive endemic species after coalescence
($R^2=0.18$, $p<0.05$ for glutamine and $R^2=0.08$, $p>0.05$ for citrate).}
\label{figS3}
\end{figure}

\begin{figure}[!h]
\centering
\internallinenumbers
\includegraphics[width=8cm,keepaspectratio]{figs/figS4.pdf}
\caption{\textbf{Pairwise competition of dominants with or without their cohorts.}
In the horizontal axis, we plot the relative frequency of the invasive dominant species in head-to-head
pairwise competition. In the vertical axis, we plot the same relative frequency when competition takes
place in the presence of the cohorts, i.e. during community coalescence.
$R^2=0.04$, $p>0.05$ for glutamine and $R^2=0.83$, $p<10^{-8}$ for citrate.}
\label{figS4}
\end{figure}
